---
title: "Sandbox Round 1 Initial Screen Report"
author: "Abby Coleman"
date: "`r Sys.Date()`"
geometry: margin = 2cm
output: 
  pdf_document:
    toc: TRUE
    
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{center}
  - \includegraphics[width=1in,height=1in]{Ellison Medical Institute Logo_Bronze.png}\LARGE\\}
  - \posttitle{\end{center}}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhead[C]{Sandbox Round 1}
  - \fancyhead[L]{\includegraphics[width=0.5in]{Ellison Medical Institute Logo_Bronze.png}}
  - \fancyfoot[L]{Ellison Medical Institute}
  - \setlength{\headheight}{30.10677pt}
---


```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir='..')
#TODO: load these in a function
source(here("R", "plot_plate_zfactors.R"))
source(here("R", "plot_zfactor_outliers.R"))
source(here("R", "plot_platemap.R"))
source(here("R", "print_platemap.R"))
source(here("R", "functions.R"))

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      fig.align = "center")

#resolve conflicts
conflicts_prefer(dplyr::filter(),
                  stats::gaussian(),
                  dplyr::select())
```


```{r data_load, message = FALSE}
ctrl0 <<- "DMSO"
ctrl100 <<- "R1881"
```


# Overall

```{r initial_screen_plots, results = "asis", warning = FALSE}
screen <- data %>%
  filter(assay_type == "Screen" & is.na(qc_flag))
# %>%
#   filter(as.numeric(short_mcule_lab) < 75 & as.numeric(short_mcule_lab) > 60)
#^ this allows you to see compound 272 more up close. that is the purple dot that appears to not be designated as a quantile outlier hit. but it actually is designated a quantile outlier hit, the median is just higher than the most visible purple point that you see.
screenplot_all <- plot_screen(screen,  input_title = "All Screening Experiments")
print(screenplot_all$plot)
#write table output to excel sheet
summary_tab <- screenplot_all$table 
write.xlsx(summary_tab, here("reports", "Hit Summary.xlsx"))
```

```{r summary_tab}
#get counts
#hits for both me and michael
quant <- data %>%
  filter(assay_type == "Screen" & is.na(qc_flag)) %>%
  filter(!is.na(scr_hit) | !is.na(tier)) 
#michael's hits that aren't quantile outliers
# manual_only <- data %>%
#   filter(assay_type == "Screen") %>%
#   filter(is.na(scr_hit) & !is.na(tier))
```

\newpage

`r get_n_mcules(quant)` compounds were designated hits by the quantile outlier method.

```{r}
screenplot_quant <- plot_screen(quant, "Hits Only", include_xaxis = TRUE)
print(screenplot_quant$plot)
```

```{r}
kable(screenplot_quant$table, digits = 2)
```

# Compound %CV's

```{r cvs, out.width = "70%", warning = FALSE}
for_plot <- data %>%
  select(experiment_id,
         experiment_type,
         assay_type,
         start_date,
         summary_cv,
         compound_cv_cutoff,
         coi,
         short_mcule_lab,
         exp_cv_result,
         qc_flag) %>%
  arrange(assay_type, experiment_id) %>%
  distinct() %>%
  filter(is.na(qc_flag))

cvplot <- ggplot(data = for_plot, aes(x = assay_type, y = exp_cv_result)) +
  geom_boxplot(outliers = FALSE) +
  geom_quasirandom(size = 3,
                   alpha = .2,
                   aes(color = short_mcule_lab)) +
  scale_color_viridis_d(end = .9, guide = "none") +
  geom_line(aes(x = assay_type, y = compound_cv_cutoff)) +
  theme_bw() +
  labs(
    x = "Assay Type",
    y = "Summary %CV",
    title = "All %CV's, Calculated by Compound"
  ) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    text = element_text(family = "serif")
  )
cvplot 
```

See Compound %CV Outliers.

```{r see_cv_outliers}
#screen data
cv_outliers <- data %>%
  filter(assay_type == "Screen" & is.na(qc_flag)) %>%
  filter(ctrl == "sample") %>%
  filter(exp_cv_result > compound_cv_cutoff) 
#color-blind safecolor scale (discrete)
# okabe <- c("black", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cv_outliers_raw_plot <- ggplot(data = cv_outliers, aes(x = short_mcule_lab, y = data.result, color = data.plate_id)) +
  geom_point(alpha = .4, size = 3) +
  theme_bw() +
  # scale_color_viridis_d() +
  labs(title = "Raw Signal for Compounds with Outlier %CV's",
       x = "MCULE",
       y = "Raw RLU",
       color = "Plate ID") +
  scale_color_paletteer_d("colorBlindness::paletteMartin")
  # scale_color_manual(values = okabe) 
cv_outliers_raw_plot

see_tab <- data %>%
  filter(assay_type == "Screen") %>%
  filter(ctrl == "sample") %>%
  filter(exp_cv_result > compound_cv_cutoff)  %>%
  select(short_mcule_lab, exp_cv_result, exp_mean_res, exp_sd_res) %>%
  distinct() %>%
  arrange(short_mcule_lab)
#TODO: change color scale

```

\newpage

# Excluded Plates

These plates were excluded for having Z'factor < 0.

```{r z_prime_lt_0}
excluded <- data %>%
  filter(assay_type == "Screen" & qc_flag == "plate z'factor < 0")
excluded_tab <- excluded %>%
  select(experiment_id, data.plate_id, plate_zprime) %>%
  distinct()
names(excluded_tab) <- c("Experiment ID", "Plate ID", "Plate Z'Factor")
kable(excluded_tab, digits = 2)
```

```{r plot_z_prime_outliers}
print(plot_zfactor_outliers(excluded))
```

